A highly-available move operation for replicated trees

### 2019-06-17 paper/move-op.tex
paper: Section 5, again

One final type of conflict that we have not discussed so far is multiple child nodes with the same parent and the same metadata.
For example, in a filesystem, two users could concurrently create files with the same name in the same directory.
Our algorithm does not take any measures to prevent such conflict, but simply retains both child nodes.
In practice, the filename collision would be resolved by making the filenames distict, e.g.\ by appending a replica identifier to the filenames.

@@

Isabelle/HOL can generate executable Haskell, OCaml, Scala, and Standard ML code from HOL definitions using a code generation mechanism~\cite{XXX}.
However,  care must be taken when using this mechanism, as:
\begin{enumerate}
\item
The mechanism is \emph{partial} in the sense that not all HOL definitions can be realised in an executable form.
Definitions making essential use of choice principles (such as our \isa{get\_parent} function) are obvious examples of definitions that will cause problems.
However, \emph{inductively defined} relations characterised by a series of introduction rules, such as our \isa{ancestor} relation are also problematic for the code generator as making them executable necessarily requires Prolog-like proof search and mode inference.
\item
Unsophisticated use of the code generator often leads to inefficient target language code.
This is especially true when using HOL's \isa{set} type, a type of potentially infinite mathematical sets.
Whilst convenient for theorem proving purposes, this type often leads to gross inefficiency in the extracted code.
\end{enumerate}

As a result, prior to evaluating our algorithms in a real-world setting, we first produce a series of refined definitions that are optimised for efficient execution rather than ease of theorem proving.
Within these definitions, we eschew the HOL \isa{set} type, and instead use a more efficient hash-map implementation provided by the Isabelle Collections Framework~\cite{XXX}.
Using an explicit \emph{simulation relation}, we show the correctness of these efficient algorithms taking our abstract algorithms as ground truth of the correct behaviour.
Moreover, we show that our correctness theorems, such as \isa{apply\_ops\_acyclic}, easily transfer to the more efficient implementations.

>>>>

One final type of conflict that we have not discussed so far is multiple child nodes with the same parent and the same metadata.
For example, in a filesystem, two users could concurrently create files with the same name in the same directory.
Our algorithm does not take any measures to prevent such conflict, but simply retains both child nodes.
In practice, the filename collision would be resolved by making the filenames distinct, e.g.\ by appending a replica identifier to the filenames.

@@

Isabelle/HOL can generate executable Haskell, OCaml, Scala, and Standard ML code from HOL type and constant definitions using a sophisticated code generation mechanism~\cite{XXX}.
However,  care must be taken when using this mechanism, as:
\begin{enumerate}
\item
The mechanism is \emph{partial} in the sense that not all HOL definitions can be realised in an executable form.
Definitions making essential use of choice principles (such as our \isa{get\_parent} function) are obvious examples of definitions that will cause problems.
However, \emph{inductively defined} relations characterised by a series of introduction rules, such as our \isa{ancestor} relation are also problematic for the code generator as making them executable necessarily requires Prolog-like proof search and mode inference.
\item
Unsophisticated use of the code generator often leads to inefficient target language code.
This is especially true when using HOL's \isa{set} type, a type of potentially infinite mathematical sets.
Whilst convenient for theorem proving purposes, this type often leads to gross inefficiency in the extracted code.
\end{enumerate}

As a result, prior to evaluating our algorithms in a real-world setting, we first produce a series of refined definitions that are optimised for efficient execution rather than ease of theorem proving.
Within these definitions, we eschew the HOL \isa{set} type, and instead use a more efficient hash-map implementation provided by the Isabelle Collections Framework~\cite{XXX}.
Write $('a::\{hashable\}, 'b)\ \isa{hm}$ for the type of hash-maps with key type $\isa{'a}$ and binding type $\isa{'b}$.
Note that keys have an explicit $\isa{hashable}$ constraint, meaning that this hash-map type may only be used with types of keys with an associated hashing function.
With this type, we represent trees by the type $(\isa{'n}::\{hashable\}, \isa{'m} \times \isa{'n})\ \isa{hm}$.
Intuitively, this type can be thought of as a mapping from \emph{child} nodes to \emph{parent} nodes, represented by the type-variable $\isa{'n}$, and any associated metadata (represented by the type variable $\isa{'m}$.

We introduce an explicit \emph{simulation relation} between our two representations of trees.
Specifically, we say that a hash-map $\isa{t}$ of type $(\isa{'n}, \isa{'m} \times \isa{'n})\ \isa{hm}$ simulates a set $T$ of type $(\isa{'n} \times \isa{'m} \times \isa{'n})\ \isa{set}$:
\begin{displaymath}
\isa{hm.lookup}\ c t = \isa{Some}\ (m, p) \leftrightarrow (p, m, c) \in T \quad\text{for all $c$, $m$, and $p$}
\end{displaymath}
where $hm.lookup$ is the (potentially failing) function that searches for a binding for key $\isa{c}$ in hash-map $t$.
We write $t \preceq T$ to assert that the hash-map $t$ and set $T$ are contained within the simulation relation, defined above.

Using an explicit \emph{simulation relation}, we show the correctness of these efficient algorithms taking our abstract algorithms as ground truth of the correct behaviour.
Moreover, we show that our correctness theorems, such as \isa{apply\_ops\_acyclic}, easily transfer to the more efficient implementations.
